---
title: "Glassdoor Income Inequality Analysis"
subtitle: "Exploring the relationships on how various factors such as gender, income, seniority, age, education affect income and bonus"
author: "Abigail Chen"
prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
##Cleaning the environment
rm(list=ls())

##Loading the packages
library(ggplot2)
library(tidyverse)
library(dplyr)
library(kableExtra)
library(modelsummary)
library(DataExplorer)
library(lspline)
library(huxtable)
library(ggthemes)
library(estimatr)
library(MASS)
library(forecast)
library(tidyverse)
library(caret) 
library(ggplot2)
library(forecast)
library(flexmix)
library(kableExtra)
options(scipen=999)

#Importing the data
df <- read_csv('https://raw.githubusercontent.com/abigailchristinechen/da2/main/term_project/data/Glassdoor_data.csv')
```

## Research Questions
Part 1:
How much do working women make compared to working men? Do we have analytical evidence surrounding the issue of salary inequality? What causes these inequalities? How is pay affected by various factors as seen in Glassdoor? How does seniority affect the wage? Can working women bridge the wage gap by getting more education? 

Part 2: 
Which gender receives more bonuses men or women? Does seniority play a role in the bonuses a person can receive? Does the years of studying affect the extra pay a person gets?

## Introduction

According to the Workplace Gender Equality Agency (WGEA), the gender pay gap measures the difference between the average earnings of women and men in the workforce. This is an internationally established measure of working women’s position in our  economy in comparison to working men. This project aims to explore the relationship between salary and gender based on the data from Glassdoor using regression analysis.  As well as looking at the relationship between, For pay we will be looking at the net income and the bonus pay. Additionally, other factors such as age, seniority, level of education of an individual will also be taken into account. 

## Motivation

According to the Workplace Gender Equality Agency (WGEA), the gender pay gap is influenced by a number of factors, including: conscious and unconscious discrimination and bias in hiring and pay decisions. women and men working in different industries and different jobs, with female-dominated industries and jobs attracting lower wages and high rates of part-time work for women.

Equality whether, socially, legally, mentally and even financially on the bases of gender are basic human rights, it's meant to be a equal for everyone. This is important in sending a positive message about an organization's values that are aligned with creating compensation systems that are transparent, fair and rewarding.  This can then lead to attracting better employees, improving employee retention, and building a healthy company culture.  


## Data
The dataset can be found in **Kaggle** (https://www.kaggle.com/nilimajauhari/glassdoor-analyze-gender-pay-gap).


```{r echo = FALSE, message = FALSE, include = FALSE}
head(df)
```

## Exploratory Data Analysis

### Data Munging
  
We can begin by looking at the summary of our data, and checking for NA's.  We have 1,000 observations.  There are no NA's.  Next, we can check out the structure of the dataset we are exploring, so that we can fix it to the correct data format.

###### -Check the summary of the Data

```{r , echo=FALSE, message = FALSE, include = FALSE}
summary(df)
```

###### -Check the structure of the Data

```{r, echo=FALSE, message = FALSE, include = FALSE}
s = str(df)
```

###### -Correct the data formats  
```{r, echo=FALSE, message = FALSE, include = FALSE }
#	Data cleaning and munging  
# Changing the data type to correct format
df$JobTitle = as.factor(df$JobTitle)
df$Gender = as.factor(df$Gender)
df$Education = as.factor(df$Education)
df$Dept = as.factor(df$Dept)
df$PerfEval = as.factor(df$PerfEval)
df$Seniority = as.factor(df$Seniority)
```

### Variables Selection

Below is the list of shortlisted variables chosen.

1. **NetIncome**: A new variable was created 'NetIncome'. It is equal to BasePay + Bonus. This is our dependent variable and gives the right depiction of net income of an individual compared to base pay alone.

2. **Bonus**:  This is the additional pay a person receives outside of their BasePay

3. **Gender**: Gives information regarding the NetIncome based on gender of a person, whether male or female. 

4. **Seniority**: This shows how long a person has been in the job and how it affects the net income, with values 1-5.

5. **Dept**: It shows difference in the salary of different departments, ranging from operations, management, administration, sales, and engineering.

6. **Education**: This shows how the income of an individual changes based on level of their education, whether they finished high school, college, masters or PHD.


```{r , include=FALSE}
## Creating an new variable Net Income (BasePay + Bonus)
df$NetIncome = df$BasePay + df$Bonus

```

## Descriptive Statistics

Below is the statistical summary of all the numerical variables involved in the data. For age, the youngest one is 18, and the oldest one is 65. The average age is 41. For the basepay, the minimum is \$34,208, and the maximum is \$179,726. The mean basepay is \$94,472. For the bonuses they receive, at minimum someone gets \$1,703, and at maximum someone gets \$11,293. The table below also shows the new variable created which is the Net Income, coming from the sum of the base pay and the bonus.  The mean and median for the Net income are close with the mean at \$100,939,	and the median at \$100,047.  The overall shape of the plot also looks like a normal distribution, as seen in Figure 1.


```{r ,echo=FALSE}
#Descriptive Statistics
P95 <- function(x){quantile(x,0.95,na.rm=T)}
P05 <- function(x){quantile(x,0.05,na.rm=T)}
Table1 <- datasummary ( ( Age ) +
                      ( BasePay ) +
                      ( Bonus ) +
                      ( NetIncome ) ~
                      Mean + Median + SD + Min + Max + P05 + P95 , 
                    data = df ,
                    title = 'Descriptive statistics' ) %>% 
  kable_styling ( latex_options = c ( "HOLD_position","scale_down", font_size = 2 ) )

Table1
```

# Part 1 : Income vs Various Factors
**Figure 1 - Distribution of net income**

```{r,echo=FALSE, fig.show='hide'}
ggplot( data = df, aes(x = NetIncome  )) + 
  ggtitle( "Figure 1: Distribution of Net Income" ) +
  geom_density( colour="darkcyan" ) 

```

mean of net income
```{r,echo=FALSE}
mean(df$NetIncome)
```

quantile of NetIncome
```{r,echo=FALSE}
quantile(df$NetIncome)
```


### Gender Breakdown

```{r,echo=FALSE}
table(df$Gender)
```

### Analyzing Gender vs Net Income
Here in Figure 2, The box plot shows that overall, working males have more net income compared to working females. Using the ANOVA test, we can see that there is a significant difference in net income of the two genders. However, we need to look at other factors as well.

ANOVA Results
```{r ,echo=FALSE}

ANOVANetIncomeGender <- aov(NetIncome ~ Gender, data = df)

summary(ANOVANetIncomeGender)

```

**Figure 2 - Gender vs NetIncome**

```{r ,echo=FALSE, fig.show='hide'}
ggplot( data = df, aes(x = Gender ,y = NetIncome, color = Gender  )) +
    ggtitle( "Figure 2: Gender vs Net Income" ) +
  geom_boxplot( )
```


### Analyzing Age vs Net Income

**Figure 3 - Age vs Net Income**
Here, in Figure 3, both the scatter plot and pearson correlation shows that age doesn't seem to play a significant role when it comes to a person's net income.

```{r,echo=FALSE, fig.show='hide'}

ggplot( data = df, aes(x = Age ,y = NetIncome,color = Gender  )) + 
  ggtitle( "Figure 3: Age vs Net Income" ) +
  geom_point( )
```

```{r,echo=FALSE }
# Trying Pearson Correlation in addition to the scatterplot
# https://libguides.library.kent.edu/spss/pearsoncorr
cor(df$NetIncome,df$Age)
```


### Analyzing Education vs Net Income
Here in Figure 4, we can see that someone with a higher education level has higher the net income, as shown in the boxplot and the ANOVA.  Looking at Figure 5, we can see that gender, education and net income analyses show that for any level of education Male have higher net income compared to the females.

**Figure 4 - Education vs Net Income**
```{r,echo=FALSE, fig.show='hide'}
ggplot( data = df, aes(x = Education ,y = NetIncome, color = Education)) + 
  ggtitle('Figure 4: Education vs. Net Income') +
  geom_boxplot()
```

ANOVA Results

```{r,echo=FALSE }
# anova results conclude that level of education plays role in deciding the net income

ANOVANetIncomeEducation <- aov(NetIncome ~ Education, data = df)

summary(ANOVANetIncomeEducation)
```

**Figure 5 - Education vs Net Income vs Gender**

```{r,echo=FALSE, fig.show='hide'}
ggplot( data = df, aes(x = Education ,y = NetIncome,color = Gender  )) + 
  ggtitle('Figure 5: Education vs Net Income vs Gender') +
  geom_boxplot( )
```


### Analyzing Department vs Net Income
The Figure 6 box plot shows that there's some slight differences in salaries for different departments, while the anova results show that different departments have different pay. While in Figure 7, we can see that in all the departments Male are paid more than females. 

**Figure 6 - Department vs Net Income**
```{r ,echo=FALSE, fig.show='hide'}
ggplot( data = df, aes(x = Dept ,y = NetIncome, color = Dept)) + 
  ggtitle('Figure 6: Department vs Net Income') +
  geom_boxplot( )
```

```{r ,echo=FALSE}
ANOVANetIncomeDept <- aov(NetIncome ~ Dept, data = df)

summary(ANOVANetIncomeDept)
```


**Figure 7 - Department vs Net Income vs Gender**

```{r,echo=FALSE, fig.show='hide'}
ggplot( data = df, aes(x = Dept ,y = NetIncome,color = Gender  )) + 
  ggtitle('Figure 7: Department vs Net Income vs Gender') +
  geom_boxplot( )
```

### Analyzing Seniority vs Net Income
In Figure 8, we'll see that the higher the seniority the higher the pay, as also concluded with the ANOVA result.s
ANOVA Results.  Figure 9 box plot shows that being more senior means more salary in all the seniority levels and males are paid more than females.

```{r ,echo=FALSE}
ANOVANetIncomeSeniority <- aov(NetIncome ~ Seniority, data = df)

summary(ANOVANetIncomeSeniority)
```

**Figure 8 - Seniority vs NetIncome**
```{r ,echo=FALSE, fig.show='hide'}
ggplot( data = df, aes(x = Seniority ,y = NetIncome, color = Seniority)) + 
  ggtitle('Figure 8: Seniority vs Net Income') +
  geom_boxplot( )
```

**Figure 9 - Seniority vs NetIncome vs Gender**
```{r ,echo=FALSE, fig.show='hide'}
ggplot( data = df, aes(x = Seniority ,y = NetIncome,color = Gender  )) + 
  ggtitle('Figure 9: Seniority vs Net Income vs Gender')
  geom_boxplot( )
```


```{r ,echo=FALSE}
## Train Test data division 80-20
set.seed(100)
n = nrow(df)
trainIndex = sample(1:n, size = round(0.8*n), replace=FALSE)
train = df[trainIndex ,]
test = df[-trainIndex ,]
```


### Regression Models

#### Formulas for four linear regression models

$$NetIncome=\beta_0+\beta_1(Gender) $$
$$NetIncome=\beta_0+\beta_1(Gender)+\beta_2(Education) $$
$$NetIncome=\beta_0+\beta_1(Gender)+\beta_2(Education)+\beta_3(Dept) $$
$$NetIncome=\beta_0+\beta_1(Gender)+\beta_2(Education)+\beta_3(Dept)+\beta_4(Seniority) $$
```{r ,echo=FALSE}
Formula1 = as.formula(NetIncome~Gender)
Formula2 = as.formula(NetIncome~Gender+Education)
Formula3 = as.formula(NetIncome~Gender+Education+Dept)
Formula4 = as.formula(NetIncome~Gender+Education+Dept+Seniority)
```



1. **Model_1**: Shows relative change on Net Income with a change in Gender. 

Summary of the model shows that the average salary for female is estimated to be 95696, whereas males are estimated a total of 95696 + 9647 = 105343. The p-value for the gender Male is very significant, suggesting that there is a statistical evidence of a difference in average salary between the genders.

```{r,echo=FALSE }
## Model Training
Model1 = lm(Formula1,train)
summary(Model1)

```


2. **Model_2**: shows relative change on Net Income with a change in Gender as well as Education. 

Summary of the model shows that the Education level of 'High School' earns on average  -1698 less,
'Masters' 5383 more and 'PhD' 8480 more. While gender still showing significant difference of pay. 

```{r,echo=FALSE }
## Model Training

Model2 = lm(Formula2,train)
summary(Model2)
```

3. **Model_3**: Shows significant effect on Net Income, with a change in Gender as well as Education and Department.

```{r,echo=FALSE }
## Model Training

Model3 = lm(Formula3,train)
summary(Model3)
```

4. **Model_4**: For this model, Gender, Education, Department along with Seniority have been considered. It shows that Seniority2 has on average 8252.0 more salary and Seniority5 has on average 37345.0 more. 



```{r,echo=FALSE }
## Model Training
Model4 = lm(Formula4,train)
summary(Model4)
```


```{r,echo=FALSE }
## On Train data

train$PredictM1 = predict(Model1,train)
ACC1 = accuracy(train$PredictM1,train$NetIncome)

train$PredictM2 = predict(Model1,train)
ACC2 = accuracy(train$PredictM2,train$NetIncome)

train$PredictM3 = predict(Model3,train)
ACC3 = accuracy(train$PredictM3,train$NetIncome)

train$PredictM4 = predict(Model4,train)
ACC4 = accuracy(train$PredictM4,train$NetIncome)

TrainDataAccuracyResults = data.frame(Model = c("Model 1","Model 2","Model 3","Model 4"),
                     MPE = c(ACC1[4],ACC2[4],ACC3[4],ACC4[4]),
                     MAPE = c(ACC1[5],ACC2[5],ACC3[5],ACC4[5])
                     )

kable(TrainDataAccuracyResults, 
      "latex", booktabs = TRUE, 
      caption = 'Accuracy results of the models on train data') %>% kable_styling(latex_options = c("hold_position","scale_down"), font_size = 2)

```

```{r ,echo=FALSE}

## Model performance on test data

test$PredictM1 = predict(Model1,test)
ACC1 = accuracy(test$PredictM1,test$NetIncome)

test$PredictM2 = predict(Model1,test)
ACC2 = accuracy(test$PredictM2,test$NetIncome)

test$PredictM3 = predict(Model3,test)
ACC3 = accuracy(test$PredictM3,test$NetIncome)

test$PredictM4 = predict(Model4,test)
ACC4 = accuracy(test$PredictM4,test$NetIncome)

TestDataAccuracyResults = data.frame(Model = c("Model 1","Model 2","Model 3","Model 4"),
                                      MPE = c(ACC1[4],ACC2[4],ACC3[4],ACC4[4]),
                                      MAPE = c(ACC1[5],ACC2[5],ACC3[5],ACC4[5])
)

kable(TestDataAccuracyResults, 
      "latex", booktabs = TRUE, 
      caption = 'Accuracy results of the models on test data') %>% kable_styling(latex_options = c("hold_position","scale_down"), font_size = 2)

```

## Conclusion : for Part 1
In conclusion, we can say that gender influences the net income of a person.

# Part 2 : Bonus vs Various Factors
**Figure 10 - Distribution of Bonus**

```{r,echo=FALSE, fig.show='hide'}
ggplot( data = df, aes(x = Bonus  )) + 
  ggtitle( "Figure 10: Distribution of Bonus" ) +
  geom_density( colour="darkcyan" ) 
```

mean of Bonus
```{r,echo=FALSE}
mean(df$Bonus)
```
quantile of Bonus
```{r,echo=FALSE}
quantile(df$Bonus)
```

### Analyzing Gender vs Bonus
Here in Figure 11, The box plot shows that overall, males and females receive almost equal bonus on average. In the ANOVA result, it clearly shows that there is no significant difference in bonus of the two genders. 

ANOVA Results
```{r ,echo=FALSE}
ANOVABonusGender <- aov(Bonus ~ Gender, data = df)
summary(ANOVABonusGender)
```

**Figure 11 - Gender vs Bonus**

```{r ,echo=FALSE, fig.show='hide'}
ggplot( data = df, aes(x = Gender ,y = Bonus, color = Gender  )) +
    ggtitle( "Figure 11: Gender vs Bonus" ) +
  geom_boxplot( )
```


### Analyzing Age vs Bonus
**Figure 12 - Age vs Bonus**
Here, in Figure 12, both the scatter plot and pearson correlation shows that age doesn't seem to play a significant role when it comes to a person's bonus.

```{r,echo=FALSE, fig.show='hide'}
ggplot( data = df, aes(x = Age ,y = Bonus,color = Gender  )) + 
  ggtitle( "Figure 12: Age vs Bonus" ) +
  geom_point( )
```

```{r,echo=FALSE }
# Trying Pearson Correlation in addition to the scatterplot
# https://libguides.library.kent.edu/spss/pearsoncorr
cor(df$Bonus,df$Age)
```


### Analyzing Education vs Bonus
Here in Figure 13, we can see that someone with a higher education level has higher the net income, as shown in the boxplot and the ANOVA.  Looking at Figure 14, we can see that for High School Education females receive slightly higher bonuses, but for college, Masters and PhD education it is almost the same regardless of gender.

**Figure 13 - Education vs Bonus**
```{r,echo=FALSE, fig.show='hide'}
ggplot( data = df, aes(x = Education ,y = Bonus, color = Education)) + 
  ggtitle('Figure 13: Education vs. Bonus') +
  geom_boxplot()
```

ANOVA Results

```{r,echo=FALSE }
ANOVABonusEducation <- aov(Bonus ~ Education, data = df)

summary(ANOVABonusEducation)
```

**Figure 14 - Education vs Bonus vs Gender**

```{r,echo=FALSE, fig.show='hide'}
ggplot( data = df, aes(x = Education ,y = Bonus,color = Gender  )) + 
  ggtitle('Figure 14: Education vs Bonus vs Gender') +
  geom_boxplot( )
```

### Analyzing Department vs Bonus
The Figure 15 box plot shows that there's some slight differences in salaries for different departments, while the anova results show that different departments have different pay. While in Figure 16, we can see that in sales and operations females receive more bonuses. While for engineering, administration and management males receive more bonuses.

## Box plot shows that there is minor difference in bounus of different departments 

**Figure 15 - Department vs Bonus**
```{r ,echo=FALSE, fig.show='hide'}
ggplot( data = df, aes(x = Dept ,y = Bonus, color = Dept)) + 
  ggtitle('Figure 15: Department vs Bonus') +
  geom_boxplot( )
```

```{r ,echo=FALSE}
ANOVABonusDept <- aov(Bonus ~ Dept, data = df)

summary(ANOVABonusDept)
```

**Figure 16 - Department vs Bonus vs Gender**

```{r,echo=FALSE, fig.show='hide'}
ggplot( data = df, aes(x = Dept ,y = Bonus,color = Gender  )) + 
  ggtitle('Figure 16: Department vs Bonus vs Gender') +
  geom_boxplot( )
```


### Analyzing Seniority vs Bonus
In Figure 17, we'll see that the higher the seniority the higher the bonus, as also concluded with the ANOVA results
ANOVA Results.  Figure 18 box plot shows that being more senior means more bonus.  In the various seniority levels, both genders have relatively equal bonuses. 

```{r ,echo=FALSE}
ANOVABonusSeniority <- aov(Bonus ~ Seniority, data = df)

summary(ANOVABonusSeniority)
```

**Figure 17 - Seniority vs Bonus**
```{r ,echo=FALSE, fig.show='hide'}
ggplot( data = df, aes(x = Seniority ,y = Bonus, color = Seniority )) + 
  ggtitle('Figure 17: Seniority vs Bonus') +
  geom_boxplot( )
```

**Figure 18 - Seniority vs Bonus vs Gender**
```{r ,echo=FALSE, fig.show='hide'}
ggplot( data = df, aes(x = Seniority ,y = Bonus, color = Gender  )) + 
  ggtitle('Figure 18: Seniority vs Bonus vs Gender') +
  geom_boxplot( )
```

```{r ,echo=FALSE}
## Train Test data division 80-20
set.seed(100)
n = nrow(df)
trainIndex = sample(1:n, size = round(0.8*n), replace=FALSE)
train = df[trainIndex ,]
test = df[-trainIndex ,]
```


### Regression Models

$$bonus=\beta_0+\beta_1(Gender) $$
$$bonus=\beta_0+\beta_1(Gender)+\beta_2(Seniority) $$
$$bonus=\beta_0+\beta_1(Gender)+\beta_2(Seniority)+\beta_3(Dept) $$
$$bonus=\beta_0+\beta_1(Gender)+\beta_2(Seniority)+\beta_3(Dept)+\beta_4(Education) $$
```{r ,echo=FALSE}
Formula5 = as.formula(Bonus~Gender)
Formula6 = as.formula(Bonus~Gender+Seniority)
Formula7 = as.formula(Bonus~Gender+Education+Dept)
Formula8 = as.formula(Bonus~Gender+Seniority+Dept+Education)
```

1. **Model_5**: shows relative change in bonus with a change in Gender. 

Summary of the model shows that the average bonus for males is estimated to be -61.33 lower than females, whereas males are estimated a total of 95696 + 9647 = 105343. The p-value for the gender Male is very high (greater than 0.05) so its not significant. Looks like there is no statistical evidence of a difference in average bonus between the genders.

```{r,echo=FALSE }
## Model Training

Model5 = lm(Formula5,train)
summary(Model5)
```
2. **Model_6**: shows relative change in Bonus with a change in Gender as well as Seniority 

Summary of the model 6 shows that the Seniority level plays a significant role in determining the bonus

```{r,echo=FALSE }
## Model Training

Model6 = lm(Formula6,train)
summary(Model6)
```

3. **Model_7**: For model 7, Gender, Seniority and Departments have been considered. The summary of the model shows that department and education have significant effect on the bonus.  


```{r,echo=FALSE }
## Model Training

Model7 = lm(Formula7,train)
summary(Model7)
```

4. **Model_8**: For this model, Gender, Education, Departments and Seniority have been considered. 

```{r,echo=FALSE }
## Model Training
Model8 = lm(Formula8,train)
summary(Model8)
```

```{r ,echo=FALSE}
## Model performance on train data

train$PredictM5 = predict(Model5,train)
ACC5 = accuracy(train$PredictM5,train$Bonus)

train$PredictM6 = predict(Model6,train)
ACC6 = accuracy(train$PredictM6,train$Bonus)

train$PredictM7 = predict(Model7,train)
ACC7 = accuracy(train$PredictM7,train$Bonus)

train$PredictM8 = predict(Model8,train)
ACC8 = accuracy(train$PredictM8,train$Bonus)

TrainDataAccuracyResults2 = data.frame(Model = c("Model 5","Model 6","Model 7","Model 8"),
                     MPE = c(ACC5[4],ACC6[4],ACC7[4],ACC8[4]),
                     MAPE = c(ACC5[5],ACC6[5],ACC7[5],ACC8[5])
                     )

kable(TrainDataAccuracyResults2, 
      "latex", booktabs = TRUE, 
      caption = 'Results of the models on train data') %>% kable_styling(latex_options = c("hold_position","scale_down"), font_size = 2)

```



```{r ,echo=FALSE}
## Model performance on test data

test$PredictM5 = predict(Model5,test)
ACC5 = accuracy(test$PredictM5,test$Bonus)

test$PredictM6 = predict(Model6,test)
ACC6 = accuracy(test$PredictM6,test$Bonus)

test$PredictM7 = predict(Model7,test)
ACC7 = accuracy(test$PredictM7,test$Bonus)

test$PredictM8 = predict(Model8,test)
ACC8 = accuracy(test$PredictM8,test$Bonus)

TestDataAccuracyResults2 = data.frame(Model = c("Model 5","Model 6","Model 7","Model 8"),
                                      MPE = c(ACC5[4],ACC6[4],ACC7[4],ACC8[4]),
                                      MAPE = c(ACC5[5],ACC6[5],ACC7[5],ACC8[5])
)

kable(TestDataAccuracyResults2, 
      "latex", booktabs = TRUE, 
      caption = 'Results of the models on test data') %>% kable_styling(latex_options = c("hold_position","scale_down"), font_size = 2)

```

#### Alternative model rlm : lm robust

```{r,echo=FALSE }

Model5R = rlm(Formula5,train)
```


```{r,echo=FALSE }

Model6R = rlm(Formula6,train)
```


```{r,echo=FALSE }

Model7R = rlm(Formula7,train)
```


```{r,echo=FALSE }
Model8R = rlm(Formula8,train)
```


```{r,echo=FALSE }
## LM Robust Model performance on train data

train$PredictM5R = predict(Model5R,train)
ACC5R = accuracy(train$PredictM5R,train$Bonus)

train$PredictM6R = predict(Model6R,train)
ACC6R = accuracy(train$PredictM6R,train$Bonus)

train$PredictM7R = predict(Model7R,train)
ACC7R = accuracy(train$PredictM7R,train$Bonus)

train$PredictM8R = predict(Model8R,train)
ACC8R = accuracy(train$PredictM8R,train$Bonus)

TrainDataAccuracyResultsLR2 = data.frame(Model = c("Model 5","Model 6","Model 7","Model 8"),
                                      MPE = c(ACC5R[4],ACC6R[4],ACC7R[4],ACC8R[4]),
                                      MAPE = c(ACC5R[5],ACC6R[5],ACC7R[5],ACC8R[5])
)
# TrainDataAccuracyResultsLR2
kable(TrainDataAccuracyResultsLR2, 
      "latex", booktabs = TRUE, 
      caption = 'Results of the alternative models (lm robust) on train data') %>% kable_styling(latex_options = c("hold_position","scale_down"), font_size = 2)

```


```{r,echo=FALSE }
## LM Robust Model performance on test data

## Model performance on test data

test$PredictM5R = predict(Model5R,test)
ACC5R = accuracy(test$PredictM5R,test$Bonus)

test$PredictM6R = predict(Model6R,test)
ACC6R = accuracy(test$PredictM6R,test$Bonus)

test$PredictM7R = predict(Model7R,test)
ACC7R = accuracy(test$PredictM7R,test$Bonus)

test$PredictM8R = predict(Model8R,test)
ACC8R = accuracy(test$PredictM8R,test$Bonus)

TestDataAccuracyResultsLR2 = data.frame(Model = c("Model 5","Model 6","Model 7","Model 8"),
                                     MPE = c(ACC5R[4],ACC6R[4],ACC7R[4],ACC8R[4]),
                                     MAPE = c(ACC5R[5],ACC6R[5],ACC7R[5],ACC8R[5])
)

# TestDataAccuracyResultsLR

kable(TestDataAccuracyResultsLR2, 
      "latex", booktabs = TRUE, 
      caption = 'Results of the alternative models (lm robust) on test data') %>% kable_styling(latex_options = c("hold_position","scale_down"), font_size = 2)

```

## Alternative model: lm robust

In robust lm, observations are not treated equally. The weight of each point is adjusted in an iterative process. rlm is less sensitive to outliers, as outliers will get reduced weight compared to simple lm which only tries to reduce the error.

## Conclusion : Part 2

In conclusion, we can say that gender influences bonus of a person. rlm didnt show much differnce mix performance good in some bad in others ... 







## Appendix

Summary of Data
```{r , echo=FALSE}
summary(df)
```

Data Format
```{r, echo=FALSE, message = FALSE}
s = str(df)
```


# Part 1: Income vs Various Factors

```{r,echo=FALSE}
# Net Income Distribution 
ggplot( data = df, aes(x = NetIncome  )) + 
   ggtitle( "Figure 1: Net Income Distribution" ) +
  geom_density( colour="darkcyan" ) 


```

```{r ,echo=FALSE}
#Gender vs Net Income Boxplot
ggplot( data = df, aes(x = Gender ,y = NetIncome, color = Gender  )) +
    ggtitle( "Figure 2: Gender vs Net Income" ) +
  geom_boxplot( )
```

```{r,echo=FALSE}
# Age vs Net Income Scatterplot
ggplot( data = df, aes(x = Age ,y = NetIncome,color = Gender  )) + 
  ggtitle( "Figure 3: Age vs Net Income" ) +
  geom_point( )
```

```{r,echo=FALSE}
ggplot( data = df, aes(x = Education ,y = NetIncome, color = Education)) + 
  ggtitle('Figure 4: Education vs. Net Income') +
  geom_boxplot()
```

```{r,echo=FALSE }
ggplot( data = df, aes(x = Education ,y = NetIncome,color = Gender  )) + 
  ggtitle('Figure 5Education vs Net Income vs Gender') +
  geom_boxplot( )
```

```{r ,echo=FALSE}
ggplot( data = df, aes(x = Dept ,y = NetIncome, color = Dept)) + 
  ggtitle('Figure 6: Department vs Net Income') +
  geom_boxplot( )
```

```{r,echo=FALSE }
ggplot( data = df, aes(x = Dept ,y = NetIncome,color = Gender  )) + 
  ggtitle('Figure 7: Department vs Net Income vs Gender') +
  geom_boxplot( )
```

```{r ,echo=FALSE}
ggplot( data = df, aes(x = Seniority ,y = NetIncome, color = Seniority)) + 
  ggtitle('Figure 8: Seniority vs Net Income') +
  geom_boxplot( )
```

```{r ,echo=FALSE}
ggplot( data = df, aes(x = Seniority ,y = NetIncome,color = Gender  )) + 
  ggtitle('Figure 9: Seniority vs Net Income vs Gender')
  geom_boxplot( )
```

# Part 2: Bonus vs Other Factors
```{r,echo=FALSE}
ggplot( data = df, aes(x = Bonus  )) + 
  ggtitle( "Figure 10: Distribution of Bonus" ) +
  geom_density( colour="darkcyan" ) 
```

```{r ,echo=FALSE}
ggplot( data = df, aes(x = Gender ,y = Bonus, color = Gender  )) +
    ggtitle( "Figure 11: Gender vs Net Bonus" ) +
  geom_boxplot( )
```

```{r,echo=FALSE}
ggplot( data = df, aes(x = Age ,y = Bonus,color = Gender  )) + 
  ggtitle( "Figure 12: Age vs Bonus" ) +
  geom_point( )
```

```{r,echo=FALSE}
ggplot( data = df, aes(x = Education ,y = Bonus, color = Education)) + 
  ggtitle('Figure 13: Education vs. Bonus') +
  geom_boxplot()
```

```{r,echo=FALSE}
ggplot( data = df, aes(x = Education ,y = Bonus,color = Gender  )) + 
  ggtitle('Figure 14: Education vs Bonus vs Gender') +
  geom_boxplot( )
```

```{r ,echo=FALSE}
ggplot( data = df, aes(x = Dept ,y = Bonus, color = Dept)) + 
  ggtitle('Figure 15: Department vs Bonus') +
  geom_boxplot( )
```

```{r,echo=FALSE}
ggplot( data = df, aes(x = Dept ,y = Bonus,color = Gender  )) + 
  ggtitle('Figure 16: Department vs Bonus vs Gender') +
  geom_boxplot( )
```

```{r ,echo=FALSE}
ggplot( data = df, aes(x = Seniority ,y = Bonus, color = Seniority )) + 
  ggtitle('Figure 17: Seniority vs Bonus') +
  geom_boxplot( )
```

```{r ,echo=FALSE}
ggplot( data = df, aes(x = Seniority ,y = Bonus, color = Gender  )) + 
  ggtitle('Figure 18: Seniority vs Bonus vs Gender') +
  geom_boxplot( )
```