---
title: "Glassdoor Income Inequality Analysis"
subtitle: "Exploring the relationships between various factors such as gender, income, seniority, age, education"
author: "Abigail Chen"
prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
##Cleaning the environment
rm(list=ls())

##Loading the packages
library(ggplot2)
library(tidyverse)
library(dplyr)
library(kableExtra)
library(modelsummary)
library(DataExplorer)
library(lspline)
library(huxtable)
library(ggthemes)
library(estimatr)
library(forecast)
library(tidyverse)
library(caret) 
library(ggplot2)
library(forecast)
library(flexmix)
library(kableExtra)
options(scipen=999)

#Importing the data
df <- read_csv('/Users/abigailchristinechen/da2/term_project/data/Glassdoor_data.csv')
```

## Research Questions

How much do working women make compared to working men? Do we have analytical evidence surrounding the issue of salary inequality? What causes these inequalities? How is pay affected by various factors as seen in Glassdoor? How does seniority affect the wage? Can working women bridge the wage gap by getting more education?


## Introduction

According to the Workplace Gender Equality Agency (WGEA), the gender pay gap measures the difference between the average earnings of women and men in the workforce. This is an internationally established measure of working women’s position in our  economy in comparison to working men. This project aims to explore the relationship between salary and gender based on the data from Glassdoor using regression analysis.  For pay we will be looking at the net income and the bonus pay. Additionally, other factors such as age, seniority, level of education of an individual will also be taken into account. 


## Motivation

According to the Workplace Gender Equality Agency (WGEA), the gender pay gap is influenced by a number of factors, including: conscious and unconscious discrimination and bias in hiring and pay decisions. women and men working in different industries and different jobs, with female-dominated industries and jobs attracting lower wages and high rates of part-time work for women.

Equality whether, socially, legally, mentally and even financially on the bases of gender are basic human rights, it's meant to be a equal for everyone. This is important in sending a positive message about an organization's values that are aligned with creating compensation systems that are transparent, fair and rewarding.  This can then lead to attracting better employees, improving employee retention, and building a healthy company culture.  


## Data
The dataset can be found in **Kaggle**(https://www.kaggle.com/nilimajauhari/glassdoor-analyze-gender-pay-gap).


```{r echo = FALSE, message = FALSE, include = FALSE}
head(df)
```

## Exploratory Data Analysis

### Data Munging
  
We can begin by looking at the summary of our data, and checking for NA's.  We have 1,000 observations.  There are no NA's.  Next, we can check out the structure of the dataset we are exploring, so that we can fix it to the correct data format.

###### -Check the summary of the Data

```{r , echo=FALSE, message = FALSE, include = FALSE}
summary(df)
```

###### -Check the structure of the Data

```{r, echo=FALSE, message = FALSE, include = FALSE}
s = str(df)
```

###### -Correct the data formats  
```{r, echo=FALSE, message = FALSE, include = FALSE }
#	Data cleaning and munging  
# Changing the data type to correct format
df$JobTitle = as.factor(df$JobTitle)
df$Gender = as.factor(df$Gender)
df$Education = as.factor(df$Education)
df$Dept = as.factor(df$Dept)
df$PerfEval = as.factor(df$PerfEval)
df$Seniority = as.factor(df$Seniority)
```

### Variables Selection

Below is the list of shortlisted variables chosen.

1. **NetIncome**: A new variable was created 'NetIncome'. It is equal to BasePay + Bonus. This is our dependent variable and gives the right depiction of net income of an individual compared to base pay alone.

2. **Gender**: Gives information regarding the NetIncome based on gender of a person, whether male or female. 

3. **Seniority**: This shows how long a person has been in the job and how it affects the net income, with values 1-5.

4. **Dept**: It shows difference in the salary of different departments, ranging from operations, management, administration, sales, and engineering.

5. **Education**: This shows how the income of an individual changes based on level of their education, whether they finished high school, college, masters or PHD.


```{r , include=FALSE}
## Creating an new variable Net Income (BasePay + Bonus)
df$NetIncome = df$BasePay + df$Bonus

```

## Descriptive Statistics

Below is the statistical summary of all the numerical variables involved in the data. For age, the youngest one is 18, and the oldest one is 65. The average age is 41. For the basepay, the minimum is \$34,208, and the maximum is \$179,726. The mean basepay is \$94,472. For the bonuses they receive, at minimum someone gets \$1,703, and at maximum someone gets \$11,293. The table below also shows the new variable created which is the Net Income, coming from the sum of the base pay and the bonus.  The mean and median for the Net income are close with the mean at \$100,939,	and the median at \$100,047.  The overall shape of the plot also looks like a normal distribution, as seen in Figure 1.


```{r ,echo=FALSE}
#Descriptive Statistics
P95 <- function(x){quantile(x,0.95,na.rm=T)}
P05 <- function(x){quantile(x,0.05,na.rm=T)}
Table1 <- datasummary ( ( Age ) +
                      ( BasePay ) +
                      ( Bonus ) +
                      ( NetIncome ) ~
                      Mean + Median + SD + Min + Max + P05 + P95 , 
                    data = df ,
                    title = 'Descriptive statistics' ) %>% 
  kable_styling ( latex_options = c ( "HOLD_position","scale_down", font_size = 2 ) )

Table1
```


**Figure 1 - Distribution of net income**

```{r,echo=FALSE, fig.show='hide'}
ggplot( data = df, aes(x = NetIncome  )) + 
  ggtitle( "Distribution of Net Income" ) +
  geom_density( colour="darkcyan" ) 

```

mean of net income
```{r,echo=FALSE}
mean(df$NetIncome)
```

quantile of NetIncome
```{r,echo=FALSE}
quantile(df$NetIncome)
```


### Gender Breakdown

```{r,echo=FALSE}
table(df$Gender)
```

### Analyzing Gender with respect to Net Income
Here in Figure 2, The box plot shows that overall, working males have more net income compared to working females. Using the ANOVA test, we can see that there is a significant difference in net income of the two genders. However, we need to look at other factors as well.

ANOVA Results
```{r ,echo=FALSE}

ANOVANetIncomeGender <- aov(NetIncome ~ Gender, data = df)

summary(ANOVANetIncomeGender)

```

**Figure 2 - Gender vs NetIncome**

```{r ,echo=FALSE, fig.show='hide'}
ggplot( data = df, aes(x = Gender ,y = NetIncome, color = Gender  )) +
    ggtitle( "Gender vs Net Income" ) +
  geom_boxplot( )
```

### Analyzing Age with respect to Net Income


**Figure 3 - Age vs Net Income**
Here, in Figure 3, both the scatter plot and pearson correlation shows that age doesn't seem to play a significant role when it comes to a person's net income.

```{r,echo=FALSE, fig.show='hide'}

ggplot( data = df, aes(x = Age ,y = NetIncome,color = Gender  )) + 
  ggtitle( "Age vs Net Income" ) +
  geom_point( )
```

```{r,echo=FALSE }
# Trying Pearson Correlation in addition to the scatterplot
# https://libguides.library.kent.edu/spss/pearsoncorr
cor(df$NetIncome,df$Age)
```

### Analyzing Education vs Net Income
Here in Figure 4, we can see that someone with a higher education level has higher the net income, as shown in the boxplot and the ANOVA.  Looking at Figure 5, we can see that gender, education and net income analyses show that for any level of education Male have higher net income compared to the females.

**Figure 4 - Education vs Net Income**
```{r,echo=FALSE, fig.show='hide'}
ggplot( data = df, aes(x = Education ,y = NetIncome, color = Education)) + 
  ggtitle('Education vs. Net Income') +
  geom_boxplot()
```

ANOVA Results

```{r,echo=FALSE }
# anova results conclude that level of education plays role in deciding the net income

ANOVANetIncomeEducation <- aov(NetIncome ~ Education, data = df)

summary(ANOVANetIncomeEducation)
```

**Figure 5 - Education vs Net Income vs Gender**

```{r,echo=FALSE, fig.show='hide'}
ggplot( data = df, aes(x = Education ,y = NetIncome,color = Gender  )) + 
  ggtitle('Education vs Net Income vs Gender') +
  geom_boxplot( )
```


### Analyzing Department vs Net Income
The Figure 6 box plot shows that there's some slight differences in salaries for different departments, while the anova results show that different departments have different pay. While in Figure 7, we can see that in all the departments Male are paid more than females. 

**Figure 6 - Department vs Net Income**
```{r ,echo=FALSE, fig.show='hide'}
ggplot( data = df, aes(x = Dept ,y = NetIncome, color = Dept)) + 
  ggtitle('Department vs Net Income') +
  geom_boxplot( )
```

```{r ,echo=FALSE}
ANOVANetIncomeDept <- aov(NetIncome ~ Dept, data = df)

summary(ANOVANetIncomeDept)
```


**Figure 7 - Department vs Net Income vs Gender**

```{r,echo=FALSE, fig.show='hide'}
ggplot( data = df, aes(x = Dept ,y = NetIncome,color = Gender  )) + 
  ggtitle('Department vs Net Income vs Gender') +
  geom_boxplot( )
```

### Analyzing Seniority vs Net Income
In Figure 8, we'll see that the higher the seniority the higher the pay, as also concluded with the ANOVA result.s
ANOVA Results.  Figure 9 box plot shows that being more senior means more salary in all the seniority levels and males are paid more than females.

```{r ,echo=FALSE}
ANOVANetIncomeSeniority <- aov(NetIncome ~ Seniority, data = df)

summary(ANOVANetIncomeSeniority)
```

**Figure 8 - Seniority vs NetIncome**
```{r ,echo=FALSE, fig.show='hide'}
ggplot( data = df, aes(x = Seniority ,y = NetIncome, color = Seniority)) + 
  ggtitle('Seniority vs Net Income') +
  geom_boxplot( )
```

**Figure 9 - Seniority vs NetIncome**
```{r ,echo=FALSE, fig.show='hide'}
ggplot( data = df, aes(x = Seniority ,y = NetIncome,color = Gender  )) + 
  ggtitle('Seniority vs Net Income vs Gender')
  geom_boxplot( )
```


```{r ,echo=FALSE}
## Train Test data division 80-20
set.seed(100)
n = nrow(df)
trainIndex = sample(1:n, size = round(0.8*n), replace=FALSE)
train = df[trainIndex ,]
test = df[-trainIndex ,]
```


### Regression Models

#### Formulas for four linear regression models

$$NetIncome=\beta_0+\beta_1(Gender) $$
$$NetIncome=\beta_0+\beta_1(Gender)+\beta_2(Education) $$
$$NetIncome=\beta_0+\beta_1(Gender)+\beta_2(Education)+\beta_3(Dept) $$
$$NetIncome=\beta_0+\beta_1(Gender)+\beta_2(Education)+\beta_3(Dept)+\beta_4(Seniority) $$
```{r ,echo=FALSE}
Formula1 = as.formula(NetIncome~Gender)
Formula2 = as.formula(NetIncome~Gender+Education)
Formula3 = as.formula(NetIncome~Gender+Education+Dept)
Formula4 = as.formula(NetIncome~Gender+Education+Dept+Seniority)
```












## Appendix
Summary of Data
```{r , echo=FALSE}
summary(df)
```

Data Format
```{r, echo=FALSE, message = FALSE}
s = str(df)
```

```{r,echo=FALSE}
# Net Income Distribution 
ggplot( data = df, aes(x = NetIncome  )) + 
   ggtitle( "Figure 1: Net Income Distribution" ) +
  geom_density( colour="darkcyan" ) 


```

```{r ,echo=FALSE}
#Gender vs Net Income Boxplot
ggplot( data = df, aes(x = Gender ,y = NetIncome, color = Gender  )) +
    ggtitle( "Figure 2: Gender vs Net Income" ) +
  geom_boxplot( )
```

```{r,echo=FALSE}
# Age vs Net Income Scatterplot
ggplot( data = df, aes(x = Age ,y = NetIncome,color = Gender  )) + 
  ggtitle( "Figure 3: Age vs Net Income" ) +
  geom_point( )
```

```{r,echo=FALSE}
ggplot( data = df, aes(x = Education ,y = NetIncome, color = Education)) + 
  ggtitle('Figure 4: Education vs. Net Income') +
  geom_boxplot()
```

```{r,echo=FALSE }
ggplot( data = df, aes(x = Education ,y = NetIncome,color = Gender  )) + 
  ggtitle('Figure 5Education vs Net Income vs Gender') +
  geom_boxplot( )
```

```{r ,echo=FALSE}
ggplot( data = df, aes(x = Dept ,y = NetIncome, color = Dept)) + 
  ggtitle('Figure 6: Department vs Net Income') +
  geom_boxplot( )
```

```{r,echo=FALSE }
ggplot( data = df, aes(x = Dept ,y = NetIncome,color = Gender  )) + 
  ggtitle('Figure 7: Department vs Net Income vs Gender') +
  geom_boxplot( )
```

```{r ,echo=FALSE, fig.show='hide'}
ggplot( data = df, aes(x = Seniority ,y = NetIncome, color = Seniority)) + 
  ggtitle('Figure 8: Seniority vs Net Income') +
  geom_boxplot( )
```

```{r ,echo=FALSE, fig.show='hide'}
ggplot( data = df, aes(x = Seniority ,y = NetIncome,color = Gender  )) + 
  ggtitle('Figure 9: Seniority vs Net Income vs Gender')
  geom_boxplot( )
```



