---
title: "Glassdoor Income Inequality Analysis"
subtitle: "Exploring the relationships between various factors such as gender, income, seniority, age, education"
author: "Abigail Chen"
prettydoc::html_pretty:
    theme: leonids
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
##Cleaning the environment
rm(list=ls())

##Loading the packages
library(ggplot2)
library(tidyverse)
library(dplyr)
library(kableExtra)
library(modelsummary)
library(DataExplorer)
library(lspline)
library(huxtable)
library(ggthemes)
library(estimatr)
library(forecast)
library(tidyverse)
library(caret) 
library(ggplot2)
library(forecast)
library(flexmix)
library(kableExtra)
options(scipen=999)

#Importing the data
df <- read_csv('/Users/abigailchristinechen/da2/term_project/data/Glassdoor_data.csv')
```

## Research Questions

How much do working women make compared to working men? Do we have analytical evidence surrounding the issue of salary inequality? What causes these inequalities? How is pay affected by various factors as seen in Glassdoor? How does seniority affect the wage? Can working women bridge the wage gap by getting more education?


## Introduction

According to the Workplace Gender Equality Agency (WGEA), the gender pay gap measures the difference between the average earnings of women and men in the workforce. This is an internationally established measure of working women’s position in our  economy in comparison to working men. This project aims to explore the relationship between salary and gender based on the data from Glassdoor using regression analysis.  For pay we will be looking at the net income and the bonus pay. Additionally, other factors such as age, seniority, level of education of an individual will also be taken into account. 


## Motivation

According to the Workplace Gender Equality Agency (WGEA), the gender pay gap is influenced by a number of factors, including: conscious and unconscious discrimination and bias in hiring and pay decisions. women and men working in different industries and different jobs, with female-dominated industries and jobs attracting lower wages and high rates of part-time work for women.

Equality whether, socially, legally, mentally and even financially on the bases of gender are basic human rights, it's meant to be a equal for everyone. This is important in sending a positive message about an organization's values that are aligned with creating compensation systems that are transparent, fair and rewarding.  This can then lead to attracting better employees, improving employee retention, and building a healthy company culture.  


## Data
The dataset can be found in **Kaggle**(https://www.kaggle.com/nilimajauhari/glassdoor-analyze-gender-pay-gap).


```{r echo = FALSE, message = FALSE, include = FALSE}
head(df)
```

## Exploratory Data Analysis

### Data Munging
  
We can begin by looking at the summary of our data, and checking for NA's.  We have 1,000 observations.  There are no NA's.  Next, we can check out the structure of the dataset we are exploring, so that we can fix it to the correct data format.

###### Check the summary of the Data

```{r , echo=FALSE}
summary(df)
```

###### Check the structure of the Data

```{r, echo=FALSE}
s = str(df)
```

###### Correcting the data formats  
```{r , include=FALSE}
#	Data cleaning and munging  
## changing the data type to correct format
df$JobTitle = as.factor(df$JobTitle)
df$Gender = as.factor(df$Gender)
df$Education = as.factor(df$Education)
df$Dept = as.factor(df$Dept)
df$PerfEval = as.factor(df$PerfEval)
df$Seniority = as.factor(df$Seniority)
```

### Variables Selection

Below is the list of shortlisted variables chosen.

1. **NetIncome**: A new variable was created 'NetIncome'. It is equal to BasePay + Bonus. This is our dependent variable and gives the right depiction of net income of an individual compared to base pay alone.

2. **Gender**: Gives information regarding the NetIncome based on gender of a person, whether male or female. 

3. **Seniority**: This shows how long a person has been in the job and how it affects the net income, with values 1-5.

4. **Dept**: It shows difference in the salary of different departments, ranging from operations, management, administration, sales, and engineering.

5. **Education**: This shows how the income of an individual changes based on level of their education, whether they finished high school, college, masters or PHD.


